---
title: "CART samples"
# citation-style: nature.csl
# documentclass: book
# link-citations: yes
# bibliography: 'cite.bib'
#output: html_notebook
output: html_document
---
<style>
.main-container {
  max-width: 1500px;!important
  margin-left: auto;!important
  margin-right: auto;!important
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      dpi = 300,
                      dev = "png",
                      dev.args = list(type = "cairo-png"))
library(gt23)
library(RMySQL)
library(openxlsx)
library(tidyverse)
library(VennDiagram)
library(kableExtra)
library(rmarkdown)
options(bitmapType='cairo')
display_venn <- function(x, ...){
  grid.newpage()
  venn_object <- venn.diagram(x, filename = NULL, ...)
  grid.draw(venn_object)
}
#library(dplyr)
```

```{r getData, warning=FALSE}
dbConn  <- dbConnect(MySQL(), group='specimen_management')
samples_tmp <- dbGetQuery(dbConn, 
                      'select * from gtsp where Trial="CART19_ALL" or Trial="CART19_CLL"') 
samples_tmp2 <- samples_tmp %>%   mutate(SamplePatientCode=ifelse(grepl('^xxx$',SamplePatientCode),sub('^p','CHP',Patient),SamplePatientCode))
rekk <-regexpr("\\..*$", samples_tmp2$SamplePatientCode , perl=TRUE)
rekk2 <- ifelse(rekk==-1,10000,rekk)
sub_rekk <- substring(samples_tmp2$SamplePatientCode,first=rekk2 )
samples <- samples_tmp2 %>%
  mutate(SpecimenInfo=ifelse(sub_rekk=='',SpecimenInfo,paste(SpecimenInfo,sub_rekk,sep = ':'))) %>%
  mutate(SamplePatientCode=sub("\\..*$",'',SamplePatientCode))
response_page1 <- openxlsx::read.xlsx("CART19 Patient_Response Info_12102020.xlsx", 1, detectDates = TRUE)
response_page2 <- openxlsx::read.xlsx("CART19 Patient_Response Info_12102020.xlsx", 2, detectDates = TRUE)



kk  <- samples %>% filter(Patient %in% response_page1$Patient.ID)
response_page2_bis <- response_page2 %>% mutate(SamplePatientCode=paste0(protocol,'-',ID))
kk2 <- samples %>% filter(SamplePatientCode %in% response_page2_bis$SamplePatientCode)
kk3 <- samples %>% filter(grepl('^[CPN]R',samples$MiscComments))
nobles_sampleData <- read.table('sampleData.tsv', sep = '\t', header = TRUE) 
nobles_sampleData_f <- nobles_sampleData %>%
  filter(Response_class!='')
kk4 <- samples %>% filter( SpecimenAccNum %in% nobles_sampleData_f$SpecimenAccNum)
found_samples_GTSP <- unique(c(kk$SpecimenAccNum,kk2$SpecimenAccNum,kk3$SpecimenAccNum,kk4$SpecimenAccNum))

missing_samples <- samples %>% filter(!(SpecimenAccNum %in% found_samples_GTSP))

sets <- list(
  page1 = kk$SpecimenAccNum, 
  page2 = kk2$SpecimenAccNum, 
  misc = kk3$SpecimenAccNum,
  cnobles = kk4$SpecimenAccNum
)
```


# Samples

A total of __`r samples %>% nrow()`__ samples have been sequenced 
( __`r samples %>% filter(Trial=="CART19_ALL") %>% nrow()`__ ALL and 
 __`r samples %>% filter(Trial=="CART19_CLL") %>% nrow()`__ CLL). We have response class
 data for __`r length(found_samples_GTSP)`__ samples. We still need response class data
 for __`r missing_samples %>% nrow()`__ samples.
 
# Response class
 
 Response class data has been acquired from 3 places

 1. A 2 pages xlmx files
 2. Miscellaneous field in the database
 3. Data from Chris Nobles
 
 
```{r responseVenn,out.width="90%",fig.align="center" ,fig.cap="source for response class"}

sets <- list(
  page1 = kk$SpecimenAccNum, 
  page2 = kk2$SpecimenAccNum, 
  misc = kk3$SpecimenAccNum,
  cnobles = kk4$SpecimenAccNum
)

display_venn(
  sets,
#  category.names = c("Set 1" , "Set 2 " , "Set 3", "Set 4"),
  # Circles
  lwd = 2,
  lty = 'blank',
  fill = c("#999999", "#E69F00", "#56B4E9", "#009E73"),
  # Numbers
  cex = .9,
  fontface = "italic",
  # Set names
  cat.cex = 1,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.dist = c(0.055, 0.055, 0.1, 0.1)
)
```

```{r responseVenn2,out.width="90%",fig.align="center" ,fig.cap="proportion of missing responses"}

miss_sets <- list(
  all = samples$SpecimenAccNum,
  cnobles = kk4$SpecimenAccNum
)

display_venn(
  miss_sets,
#  category.names = c("Set 1" , "Set 2 " , "Set 3", "Set 4"),
  # Circles
  lwd = 2,
  lty = 'blank',
  fill = c("#999999", "#E69F00"),
  # Numbers
  cex = .9,
  fontface = "italic",
  # Set names
  cat.cex = 1,
  cat.fontface = "bold",
  cat.default.pos = "outer"
)
```

# Missing data 

```{r missingTable, message=FALSE, warning=FALSE}
missing_table <- samples %>% filter( SpecimenAccNum %in% missing_samples$SpecimenAccNum) %>%
#  select(c(SpecimenAccNum,SamplePatientCode,Patient,Sender)) %>%
  group_by(Trial,SamplePatientCode,Patient) %>% summarise(NumberSamples=n(),.groups='drop') %>%
  arrange(Trial,Patient) 
missing_table %>%
  kable( "html", table.attr = "style = \"color: black;\"") %>%
  kable_styling("hover", full_width = F)
write.xlsx(missing_table,'CART19_missing_response_samples_20211117.xls',overwrite = TRUE)
```



# Full list missing samples

```{r missingTableF}
samples %>% filter( SpecimenAccNum %in% missing_samples$SpecimenAccNum) %>%
  select(c(SpecimenAccNum,SamplePatientCode,Patient,Sender)) %>%
  kable( "html",table.attr = "style = \"color: black;\"") %>%
  kable_styling("hover", full_width = F)
```

```{r addResponse, eval=FALSE, include=FALSE}
results <- readRDS('result_full_ALL_CLL.rds')
resp_class <- nobles_sampleData %>%
  select(c('SpecimenAccNum',"Response_class")) %>%
  mutate(specimen=SpecimenAccNum) %>%
  mutate(Response_class= as.factor( ifelse(Response_class=='',"MISS",as.character(Response_class)))) %>%
  mutate(SpecimenAccNum=NULL)

kkkk <- left_join(results,resp_class,by = 'specimen') %>%
  mutate(Response_class= as.factor( ifelse(is.na(Response_class),"MISS",as.character(Response_class))))

saveRDS(kkkk, file = 'result_full_ALL_CLL_ML.rds')  
```

# Make full patient list

```{r full patient list}
nobles_class <- nobles_sampleData %>%
  select(c('SpecimenAccNum',"Response","Response_class",'Trial_ID')) %>%
  mutate(Response_class= as.factor( ifelse(Response_class=='',"MISSING",as.character(Response_class)))) %>%
  mutate(Response= as.factor( ifelse(Response=='',"MISSING",as.character(Response)))) 

current_table  <- left_join(samples,nobles_class,by = 'SpecimenAccNum') %>%
  mutate(Response_class= as.factor( ifelse(is.na(Response_class),"MISSING",as.character(Response_class)))) %>%
  mutate(Response= as.factor( ifelse(is.na(Response),"MISSING",as.character(Response)))) %>%
  group_by(Trial,Trial_ID,SamplePatientCode,Patient,Response,Response_class) %>% summarise(NumberSamples=n(),.groups='drop')
paged_table(current_table,options = list(rows.print = 20))
#current_table
wb <- openxlsx::createWorkbook()
openxlsx::addWorksheet(wb, "All_samples")
openxlsx::writeDataTable(wb,"All_samples",current_table)
openxlsx::addWorksheet(wb, "Missing_response_samples")
openxlsx::writeDataTable(wb,"Missing_response_samples",current_table %>% filter(Response=='MISSING'))
openxlsx::saveWorkbook(wb, file = "ALL_CLL_samples_20211124.xlsx", overwrite = TRUE)
#write.xlsx(missing_table,'CART19_missing_response_samples_20211117.xls',overwrite = TRUE)
```